## Contextualism {#topten-contextualism}


:::{#contextualism-graph style="float:right; width:60%"}

```{r contextualism-graph, fig.height=4, fig.width=4.5, fig.cap=cap_text}
jjj <- 7
the_cat <- topics_big_table$cat_num[jjj]
target_articles <- round_three %>% filter(cat_num == the_cat)
active_precursors <- precursors %>% filter(cat_num == the_cat)
category_properties <- category_years %>% filter(cat_num == the_cat)
source('appendix-graphs.R') 
p1
cap_text <- paste0("Frequency of articles in ", category_properties$cat_real_name[1], " across all journals from 2000-2019")

journal_distribution <- target_articles %>%
  group_by(journal) %>%
  tally(wt = s) %>%
  arrange(-n) %>%
  mutate(n = round(n, 2)) %>%
  slice(1:10)
```
:::

<br><br>

**Articles**: `r round(category_properties$n[1],1)`

**Citations**: `r round(category_properties$c[1],1)`

**Mean Year**: `r toString(round(category_properties$y[1],2))` <br>
**Relative Age**: `r category_properties$y_rank[1]` <br>
**Age in Group 1**:  `r category_properties$class_y_rank[1]` 

**Primary Journals**

1. `r journal_distribution$journal[1]`
2. `r journal_distribution$journal[2]`
3. `r journal_distribution$journal[3]`

:::{#contextualism-papers style="clear:both"}

```{r contextualism-papers}
cat("**Most Cited Articles**\n\n")

target_articles <- round_three %>% filter(cat_num == the_cat)
cat("::: apa-reference \n\n")  
for (jj in 1:5){
  cat(target_articles$longcite[jj], " \n\n", sep="")
}
cat(":::\n\n")

cat("<br>**[Full Data](#t-Contextualism)**\n\n")

```

:::

<br>

**Comments**

Hello world



