## Reasons {#topten-reasons}

:::{#reasons-graph style="float:right; width:60%"}

```{r reasons-graph, fig.height=4, fig.width=4.5, fig.cap=cap_text}
jjj <- 55
the_cat <- topics_big_table$cat_num[jjj]
target_articles <- round_three %>% filter(cat_num == the_cat)
active_precursors <- precursors %>% filter(cat_num == the_cat)
category_properties <- category_years %>% filter(cat_num == the_cat)
source('appendix-graphs.R') 
p1
cap_text <- paste0("Frequency of articles in ", category_properties$cat_real_name[1], " across all journals from 2000-2019")

journal_distribution <- target_articles %>%
  group_by(journal) %>%
  tally(wt = s) %>%
  arrange(-n) %>%
  mutate(n = round(n, 2)) %>%
  slice(1:10)
```
:::

<br><br>

**Articles**: `r round(category_properties$n[1],1)`

**Citations**: `r round(category_properties$c[1],1)`

**Mean Year**: `r toString(round(category_properties$y[1],2))` <br>
**Relative Age**: `r category_properties$y_rank[1]` <br>
**Age in Group 1**:  `r category_properties$class_y_rank[1]` 

**Primary Journals**

1. `r journal_distribution$journal[1]`
2. `r journal_distribution$journal[2]`
3. `r journal_distribution$journal[3]`

:::{#reasons-papers style="clear:both"}

```{r reasons-papers}
cat("**Most Cited Articles**\n\n")

target_articles <- round_three %>% filter(cat_num == the_cat)
cat("::: apa-reference \n\n")  
for (jj in 1:5){
  cat(target_articles$longcite[jj], " \n\n", sep="")
}
cat(":::\n\n")

cat("<br>**[Full Data](#t-Reasons)**\n\n")

```

:::

<br>

**Comments**

Hello world

1. Modifer Theory vs Reified Theory
2. Prima facie vs pro tanto
3. Wrong Kind of Reasons
4. Reasons Possession - do evil people possess reasons to be good
5. Ethics and Epistemology
  A. Is it true that only evidence provides reasons?
  B. Is it true that only knowledge provides reason for action?
  C. Is it true that immoral action is unreasonable, in just the same sense that irrational belief is?
6. What TF are reasons
  Propositions, Facts, Mental States, something else
  I kind of don't think there is a good answer here, which is why I still like modifier theory

